---
title: "AnnualSummary_Child"
author: "John"
date: "Monday, August 25, 2014"
output: pdf_document
---


# `r .year` 

```{r totals}
# function to convert words to Title Case
capwords <- function(s) {
    cap <- function(s) paste(toupper(substring(s, 1, 1)),
                  {s <- substring(s, 2); tolower(s)},
                             sep = "", collapse = " " )
    sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}

nRefugees = ednll. %>% filter(year == .year, AlienType == 'R') %>% nrow %>% comma

refugeeCountries = ednll. %>% filter(year == .year, AlienType == 'R') %$% unique(nationalityCountry) %>% length

cubans = ednll. %>% filter(year == .year, AlienType == 'R', examCountry == 'CUBA') %>% nrow %>% comma

top4_nonCuban = ednll. %>% 
     filter(year == .year, AlienType == 'R', !( examCountry == 'CUBA')) %>% 
     group_by(nationalityCountry) %>% 
     summarize(count = n()) %>% 
     arrange(-count) %>% 
     mutate(cumtotal = cumsum(count), 
            pertotal = cumtotal/sum(count))

nTop4nonCuban = top4_nonCuban %>% .[4, 4] %>% percent
top4_nonCubanCountries = top4_nonCuban %>% .[1:4, 1] %>% paste(collapse=', ') %>% capwords

classbs = ednll. %>% filter(year == .year, AlienType == 'I') %>% nrow %>% comma

classbCountries = ednll. %>% filter(year == .year, AlienType == 'I') %$% unique(examCountry) %>% length

top5B = ednll. %>% 
     filter(year == .year, AlienType == 'I') %>% 
     group_by(Country) %>% 
     summarize(count = n()) %>% 
     arrange(-count) %>% 
     mutate(cumtotal = cumsum(count), pertotal = cumtotal/sum(count)) 

nTop5B = top5B  %>% .[4, 4] %>% percent
top5BCountries = top5B  %>% .[1:5, 1] %>% paste(collapse=', ') %>% capwords

newImmigrants = dhs %>% filter(year == .year, TypeOfLPRAdmissionStatus == 'New Arrival') %$% sum(Arrivals) %>% comma

newImmigrantCountries = dhs %>% 
     filter(year == .year, TypeOfLPRAdmissionStatus == 'New Arrival', Arrivals>0) %$% unique(Country) %>% length

nAdoptees = dhs %>% 
     filter(year == .year, TypeOfLPRAdmissionStatus == 'New Arrival', ClassOfAdmission == 'Adoptees') %$% 
     sum(Arrivals) %>% comma

AdopteeCountries = dhs %>% 
     filter(year == .year, TypeOfLPRAdmissionStatus == 'New Arrival', ClassOfAdmission == 'Adoptees') %$% 
     unique(Country) %>% length
     
nAos = dhs %>% filter(year == .year, TypeOfLPRAdmissionStatus == 'AOS') %$% sum(Arrivals) %>% comma

aosCountries = dhs %>% filter(year == .year, TypeOfLPRAdmissionStatus == 'AOS', Arrivals>0) %$% unique(Country) %>% length

```

In `r .year`, there were

+ `r nRefugees` refugee arrivals from `r refugeeCountries` nationalities
     + Of these, `r cubans`were from Cuba
     + Of the remaining, 4 countries accounted for `r nTop4nonCuban`:`r top4_nonCubanCountries `
  
+ `r classbs` Class B immigrant arrivals from `r classbCountries` countries.
     + The top 5 accounted for `r nTop5B ` : `r top5BCountries `

+ `r newImmigrants` new immigrants from `r newImmigrantCountries` countries.^[DHS, fiscal year 2013]
    + Of these, `r nAdoptees` were adoptees from `r AdopteeCountries` countries.

+ `r nAos` temporary residents from `r aosCountries` countries who adjusted status to legal permanent resident.^[DHS, fiscal year 2013]
   

```{r yearly}

yearly = arrivals %>%
     dplyr::group_by(year, AlienType) %>%
     dplyr::summarize( Arrivals = sum(Arrivals)) %>%
     mutate(
          count = Arrivals, 
          Arrivals = factor(AlienType, levels = c('New Arrival', 'AOS', 'B', 'R'),
                              labels = c('New Arrival' = 'New Immigrant Arrivals',
                                         'AOs' = 'Status Adjusters', 
                                         'B' = 'Class B New Immigrants',
                                         'R' = 'Refugees'))) %>%
     as.data.frame(yearly)  %>%   
     select( -AlienType) 

```

## Arrivals^[The demographic snapshot for immigrant arrivals is from DHS (available through 2013), and the one for refugees is from CDC.] 

```{r arrival_table, eval=FALSE}
yearly. = yearly %>%
     spread(year, count)

data = as.matrix(yearly.[, -1])
dimnames(data)[1] = yearly.[1]
data = t(data)

kable( comma(data), align = c('l', rep('r', ncol(data)-1) ))

```

```{r topArrivals}

imms = dhs %>% 
     filter(year == .year, TypeOfLPRAdmissionStatus == 'New Arrival', Arrivals>0) %>%
     mutate(DHSCountry = capwords(as.character(Country))) %>%
     group_by(Country, ISO3) %>%
     summarize( newImm = round(sum(Count, na.rm=TRUE)) ) %>%
     left_join(population, by='ISO3')

noC = dhs %>% anti_join(population, by='ISO3') %$% table(CountryOfBirth)

```

```{r arrival_chart, fig.height= 3, fig.cap= ""}
g = ggplot(data = yearly, aes(x=year, y=count)) 
g = g + geom_point(aes(color = Arrivals)) 
g = g + geom_line(aes(color = Arrivals, group=Arrivals))

g = g + scale_color_manual('', values=c('Green', 'Blue', 'Red', 'Purple' )) +
     scale_y_continuous(labels=comma) +
     xlab("Year") + 
     ylab("Arrivals")  +
     theme_bw() +
     theme(legend.position = 'top')
g
```

```{r immigrants_by_age, eval=FALSE}
a = dhs %>%
      filter(year == .year) %>%
     select(Country, year, Arrivals, Sex, agegroup, AlienType) %>%
     dplyr::group_by(year, AlienType, agegroup, Sex) %>%
     dplyr::summarize(Arrivals = sum(Arrivals)) %>%
     mutate( Arrivals = ifelse(Sex %in% 'F', -1*Arrivals, Arrivals),
             percent.male = round( 
                          sum(Arrivals * (Sex %in% "M"))*100/sum(abs(Arrivals) * (Sex %in% c("F", "M")) ) 
                          )
             ) 

v = a %>%
     dplyr::group_by(year, AlienType, agegroup) %>%
     dplyr::summarize(Arrivals = sum(abs(Arrivals)) ) %>%
     mutate( visa.percent = round(100*abs(Arrivals)/sum(Arrivals) ) ) %>%
     select(-Arrivals)
     
av = left_join(a,v, by=c('year', 'AlienType', 'agegroup'))     

```
 
```{r tornado_immigrants, fig.height = 4, fig.cap="The age and sex distribution of newly arrived immigrants, 2013", eval=FALSE}
text.scale = 4

# trans_new("abs_trans", function(x) abs(x), function(x) abs(x), format = comma, domain = c(-Inf, Inf))

data = av %>% filter(AlienType %in% 'New Arrival')
max.count=max(abs(data$Arrivals))

       ggTornado = 
       ggplot(data, aes(x=agegroup, y=Arrivals, fill=Sex)) + 
            geom_bar(stat="identity", position="identity", width = .95) + 
            
# #             # Percent total label  # show value on female side only
#             geom_text(aes(label = ifelse(Sex %in% "F" ,   
#                                          paste(" ", visa.percent, "% ", sep=""),
#                                          ""),
#                           
#                           y=Arrivals,
#                           hjust = ifelse(visa.percent>20, 0, 1),
#                           vjust=ifelse(visa.percent>33, 0, 0.5) ,
#                           color = ifelse(visa.percent>20, 
#                                          "white", "black")
#                           ),
#                       size=text.scale                         
#                       ) +
#      
#             geom_text(aes(label = ifelse(Sex %in% "F" & visa.percent>20,
#                                          " of total",
#                                          ""),
#                           y=Arrivals,
#                           hjust = ifelse(visa.percent>20, 0, -1),
#                           color = ifelse(visa.percent>20, 
#                                          "white", "black")
#                           ),
#                       size=text.scale,
#                       vjust=1 ) +
            
            # Percent male label  # show value on male side only
            geom_text(aes(
                         label = ifelse(Sex %in% "M"  , 
                                         paste(" ", percent.male, "% ", sep=""),
                                         ""
                                         ),
                         hjust = ifelse(visa.percent>33, 1, 0),
                         vjust= ifelse(visa.percent>33, 0, 0.5) ,
                         color = ifelse(visa.percent>33, 'white', 'black'),
                          y=Arrivals
                          ),
                      size= text.scale - 1
                      ) +
     
            geom_text(aes(label = ifelse(Sex %in% "M" & visa.percent>20, 
                                         " male  ",
                                         ""
                                         ),
                          hjust = ifelse(visa.percent>33, 1, 0),
                          vjust= 1,
                          color = ifelse(visa.percent>33, 'white', 'black'),
                          y=Arrivals
                          ), 
                      size = text.scale - 1
                      ) +
            
            
            scale_size_continuous(range=c(4,8), guide=FALSE) + 
            scale_color_manual(values=c("black", "white"), guide=FALSE) +
            scale_y_continuous( labels = comma
                                ) + #  comma format for populations
            scale_fill_manual(values=c("red", "blue", "grey"), guide=FALSE) + 
            xlab("Age\n") +
            ylab("\nImmigrant Arrivals, 2013") + 
            coord_flip() + 
            theme( axis.title.x = element_text(face="bold", size=14),
                   axis.title.y = element_text(face="bold", size=14),
                   axis.text.x = element_text(face="bold", size=12),
                   axis.text.y = element_text(face="bold", size=12),
                   axis.ticks = element_blank(),
                   panel.background =  element_rect(fill = NA, colour = NA), 
                   panel.border =      element_rect(fill = NA, colour=NA), 
                   panel.grid.major =  element_line(colour = NA, size = 0.2),
                   panel.grid.minor =  element_line(colour = NA, size = 0.5)
            ) 

ggTornado.age = data %>% group_by(agegroup) %>% summarize(Arrivals = sum(abs(Arrivals))) %>% 
               ggplot(aes(x=agegroup, y=Arrivals/sum(Arrivals))) + 
               geom_histogram(stat="identity", position="identity", width = .95) +
               geom_text( aes(label = percent( round( abs( Arrivals/sum(abs(Arrivals)) ), 2) ),
                              y=Arrivals/sum(abs(Arrivals))/2 ,
                              color = ifelse( abs( Arrivals/sum(abs(Arrivals)) )>.25 , "white", "black"),
                              hjust = ifelse( abs( Arrivals/sum(abs(Arrivals)) )>.25 , 0.5, -1)
                              ),
                         vjust=0.5, size = text.scale 
                         )+  
               scale_color_manual(values=c("black", "white"), guide=FALSE) +
               ylab("\n  ") + 
               xlab("\n  ") + 
               coord_flip() + 
                    theme_bw() +
                    theme( axis.title.y = element_blank(),
                           axis.text.y = element_blank() ,
                        axis.title.x = element_text(face="bold", size=14),
                        axis.text.x = element_text( face="bold", size=12, color = 'white') ,
                        axis.ticks = element_blank(),
                        panel.background =  element_rect(fill = NA, colour = NA), 
                        panel.border =      element_rect(fill = NA, colour=NA), 
                        panel.grid.major =  element_line(colour = NA, size = 0.2),
                        panel.grid.minor =  element_line(colour = NA, size = 0.5)
                 ) 
 


ggTornado.male = data %>% group_by(Sex) %>% summarize(Arrivals = sum(Arrivals)) %>% 
               filter(Sex %in% c('F', 'M')) %>%
               ggplot(aes(x= "Sex", y=Arrivals , fill = Sex)) + 
               geom_bar(stat="identity", position="identity", width = .95) +
               geom_text( aes(label = percent( round( abs( Arrivals/sum(abs(Arrivals)) ), 2) ),
                              y=Arrivals/2 ),
                         hjust = 0.5, vjust=0.5 , x = 1, color="white",
                         size = text.scale
                         ) +  
               annotate( 'text', label = 'Female', y= -max(-1*data$Arrivals) , x = 0 , 
                         color="black", hjust=0.5, vjust=0) +  
               annotate( 'text', label = 'Male', y= max(data$Arrivals) , x = 0 , 
                         color="black", hjust=0.5, vjust=0) + 
               scale_y_continuous(labels = percent) + 
               scale_fill_manual(values=c("red", "blue", "grey"), guide=FALSE) +
               ylab("\n  ") + 
               xlab("\n  ") + 
               coord_flip() +
               theme_bw() +
               theme( axis.title.x = element_blank(),
                   axis.title.y = element_text(face="bold", size=14),
                   axis.ticks = element_blank(),
                   axis.text.x = element_blank(),
                   axis.text.y =  element_text( face="bold", size=12, color = 'black'),
                   panel.background =  element_rect(fill = NA, colour = NA), 
                   panel.border =      element_rect(fill = NA, colour=NA), 
                   panel.grid.major =  element_line(colour = NA, size = 0.2),
                   panel.grid.minor =  element_line(colour = NA, size = 0.5)
            ) 

empty = grob()

grid.arrange(empty, ggTornado.male, ggTornado.age, ggTornado, ncol=2, nrow=2, 
             widths=c(2,6), heights=c(2,6))

# grid.arrange(ggTornado, ggTornado.age, ncol=2, nrow=1, widths=c(6, 1))

```

```{r refugees_by_age, eval=FALSE}
a = arrivals %>%
      filter(year == .year, AlienType %in% 'R') %>%
     select(year, Arrivals, Sex, agegroup, AlienType) %>%
     dplyr::group_by(year, AlienType, agegroup, Sex) %>%
     dplyr::summarize(Arrivals = sum(Arrivals)) %>%
     mutate( Arrivals = ifelse(Sex %in% 'F', -1*Arrivals, Arrivals),
             percent.male = round( 
                          sum(Arrivals * (Sex %in% "M"))*100/sum(abs(Arrivals)* (Sex %in% c("F", "M")) ) 
                          )
             ) 

v = a %>%
     dplyr::group_by(year, AlienType, agegroup) %>%
     dplyr::summarize(Arrivals = sum(abs(Arrivals)) ) %>%
     mutate( visa.percent = round(100*abs(Arrivals)/sum(Arrivals) ) ) %>%
     select(-Arrivals)
     
av = left_join(a,v, by=c('year', 'AlienType', 'agegroup'))     

```
 
```{r tornado_refugees, fig.height = 4, fig.cap="The age and sex distribution of refugees, 2013", eval=FALSE}
text.scale = 4

# trans_new("abs_trans", function(x) abs(x), function(x) abs(x), format = comma, domain = c(-Inf, Inf))

data = av %>% filter(AlienType %in% 'R')
max.count=max(abs(data$Arrivals))

       ggTornado = 
       ggplot(data, aes(x=agegroup, y=Arrivals, fill=Sex)) + 
            geom_bar(stat="identity", position="identity", width = .95) + 
            
# #             # Percent total label  # show value on female side only
#             geom_text(aes(label = ifelse(Sex %in% "F" ,   
#                                          paste(" ", visa.percent, "% ", sep=""),
#                                          ""),
#                           
#                           y=Arrivals,
#                           hjust = ifelse(visa.percent>20, 0, 1),
#                           vjust=ifelse(visa.percent>33, 0, 0.5) ,
#                           color = ifelse(visa.percent>20, 
#                                          "white", "black")
#                           ),
#                       size=text.scale                         
#                       ) +
#      
#             geom_text(aes(label = ifelse(Sex %in% "F" & visa.percent>20,
#                                          " of total",
#                                          ""),
#                           y=Arrivals,
#                           hjust = ifelse(visa.percent>20, 0, -1),
#                           color = ifelse(visa.percent>20, 
#                                          "white", "black")
#                           ),
#                       size=text.scale,
#                       vjust=1 ) +
            
            # Percent male label  # show value on male side only
            geom_text(aes(
                         label = ifelse(Sex %in% "M"  , 
                                         paste(" ", percent.male, "% ", sep=""),
                                         ""
                                         ),
                         hjust = ifelse(visa.percent>33, 1, 0),
                         vjust= ifelse(visa.percent>33, 0, 0.5) ,
                         color = ifelse(visa.percent>33, 'white', 'black'),
                          y=Arrivals
                          ),
                      size= text.scale - 1
                      ) +
     
            geom_text(aes(label = ifelse(Sex %in% "M" & visa.percent>20, 
                                         " male  ",
                                         ""
                                         ),
                          hjust = ifelse(visa.percent>33, 1, 0),
                          vjust= 1,
                          color = ifelse(visa.percent>33, 'white', 'black'),
                          y=Arrivals
                          ), 
                      size = text.scale - 1
                      ) +
            
            
            scale_size_continuous(range=c(4,8), guide=FALSE) + 
            scale_color_manual(values=c("black", "white"), guide=FALSE) +
            scale_y_continuous( labels = comma
                                ) + #  comma format for populations
            scale_fill_manual(values=c("red", "blue", "grey"), guide=FALSE) + 
            xlab("Age\n") +
            ylab("\nRefugees, 2013") + 
            coord_flip() + 
            theme( axis.title.x = element_text(face="bold", size=14),
                   axis.title.y = element_text(face="bold", size=14),
                   axis.text.x = element_text(face="bold", size=12),
                   axis.text.y = element_text(face="bold", size=12),
                   axis.ticks = element_blank(),
                   panel.background =  element_rect(fill = NA, colour = NA), 
                   panel.border =      element_rect(fill = NA, colour=NA), 
                   panel.grid.major =  element_line(colour = NA, size = 0.2),
                   panel.grid.minor =  element_line(colour = NA, size = 0.5)
            ) 

ggTornado.age = data %>% group_by(agegroup) %>% summarize(Arrivals = sum(abs(Arrivals))) %>% 
               ggplot(aes(x=agegroup, y=Arrivals/sum(Arrivals))) + 
               geom_histogram(stat="identity", position="identity", width = .95) +
               geom_text( aes(label = percent( round( abs( Arrivals/sum(abs(Arrivals)) ), 2) ),
                              y=Arrivals/sum(abs(Arrivals))/2 ,
                              color = ifelse( abs( Arrivals/sum(abs(Arrivals)) )>.25 , "white", "black"),
                              hjust = ifelse( abs( Arrivals/sum(abs(Arrivals)) )>.25 , 0.5, -1)
                              ),
                         vjust=0.5, size = text.scale 
                         )+  
               scale_color_manual(values=c("black", "white"), guide=FALSE) +
               ylab("\n  ") + 
               xlab("\n  ") + 
               coord_flip() + 
                    theme_bw() +
                    theme( axis.title.y = element_blank(),
                           axis.text.y = element_blank() ,
                        axis.title.x = element_text(face="bold", size=14),
                        axis.text.x = element_text( face="bold", size=12, color = 'white') ,
                        axis.ticks = element_blank(),
                        panel.background =  element_rect(fill = NA, colour = NA), 
                        panel.border =      element_rect(fill = NA, colour=NA), 
                        panel.grid.major =  element_line(colour = NA, size = 0.2),
                        panel.grid.minor =  element_line(colour = NA, size = 0.5)
                 ) 
 


ggTornado.male = data %>% group_by(Sex) %>% summarize(Arrivals = sum(Arrivals)) %>% 
               filter(Sex %in% c('F', 'M')) %>%
               ggplot(aes(x= "Sex", y=Arrivals , fill = Sex)) + 
               geom_bar(stat="identity", position="identity", width = .95) +
               geom_text( aes(label = percent( round( abs( Arrivals/sum(abs(Arrivals)) ), 2) ),
                              y=Arrivals/2 ),
                         hjust = 0.5, vjust=0.5 , x = 1, color="white",
                         size = text.scale
                         ) +  
               annotate( 'text', label = 'Female', y= -max(-1*data$Arrivals) , x = 0 , 
                         color="black", hjust=0.5, vjust=0) +  
               annotate( 'text', label = 'Male', y= max(data$Arrivals) , x = 0 , 
                         color="black", hjust=0.5, vjust=0) + 
               scale_y_continuous(labels = percent) + 
               scale_fill_manual(values=c("red", "blue", "grey"), guide=FALSE) +
               ylab("\n  ") + 
               xlab("\n  ") + 
               coord_flip() +
               theme_bw() +
               theme( axis.title.x = element_blank(),
                   axis.title.y = element_text(face="bold", size=14),
                   axis.ticks = element_blank(),
                   axis.text.x = element_blank(),
                   axis.text.y =  element_text( face="bold", size=12, color = 'black'),
                   panel.background =  element_rect(fill = NA, colour = NA), 
                   panel.border =      element_rect(fill = NA, colour=NA), 
                   panel.grid.major =  element_line(colour = NA, size = 0.2),
                   panel.grid.minor =  element_line(colour = NA, size = 0.5)
            ) 

empty = grob()

grid.arrange(empty, ggTornado.male, ggTornado.age, ggTornado, ncol=2, nrow=2, 
             widths=c(2,6), heights=c(2,6))

# grid.arrange(ggTornado, ggTornado.age, ncol=2, nrow=1, widths=c(6, 1))

```
